---
title: 'Group 3: Initial Analysis'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(EloRating)
```

## 1. Loading the data 

### Calculating Elo Score

The following code segment provided by the client will calculate the elo ratings of cows (Those only need to be ran once).

#### Sampling replacements
```{r elo score feed, message=TRUE}
all_replacement_THI <- read.csv('../data/replacement_with_THI.csv')
water <- all_replacement_THI %>% filter(Bin_type == 'Water')
feed <- all_replacement_THI %>% filter(Bin_type == 'Feed')
THI_threshold <- 65
water_high <- all_replacement_THI %>% filter(Bin_type == 'Water', THI_max>=THI_threshold)
water_low <- all_replacement_THI %>% filter(Bin_type == 'Water', THI_max<THI_threshold)
feed_high <- all_replacement_THI %>% filter(Bin_type == 'Feed', THI_max>=THI_threshold)
feed_low <- all_replacement_THI %>% filter(Bin_type == 'Feed', THI_max<THI_threshold)
# water_high_groupby <- water_high %>% group_by(Actor_cow, Reactor_cow) %>% summarise(replacements = n(), .groups = 'drop')
water_cows <- unique(water$Reactor_cow,water$Actor_cow)
feed_cows <- unique(feed$Reactor_cow,feed$Actor_cow)
intersection <- Reduce(intersect, list(water_cows,feed_cows))

water_high_cows <- unique(water_high$Reactor_cow,water_high$Actor_cow)
water_low_cows <- unique(water_low$Reactor_cow,water_low$Actor_cow)
feed_high_cows <- unique(feed_high$Reactor_cow,feed_high$Actor_cow)
feed_low_cows <-unique(feed_low$Reactor_cow,feed_low$Actor_cow)
sub_intersection <- Reduce(intersect, list(water_high_cows, water_low_cows, feed_high_cows, feed_low_cows))

water <- water %>%
  filter(Actor_cow %in% intersection & Reactor_cow %in% intersection)

feed <- feed %>%
  filter(Actor_cow %in% intersection & Reactor_cow %in% intersection)

water_high <- water_high %>%
  filter(Actor_cow %in% sub_intersection & Reactor_cow %in% sub_intersection)

water_low <- water_low %>%
  filter(Actor_cow %in% sub_intersection & Reactor_cow %in% sub_intersection)

feed_high <- feed_high %>%
  filter(Actor_cow %in% sub_intersection & Reactor_cow %in% sub_intersection)

feed_low <- feed_low %>%
  filter(Actor_cow %in% sub_intersection & Reactor_cow %in% sub_intersection)

# 
# pairs <- unique(water_high$Reactor_cow * 10000 + water_high$Actor_cow)
# 
# water_high_groupby <- water_high_groupby %>%
#   mutate(pair1 = Actor_cow * 10000 + Reactor_cow) %>%
#   mutate(pair2 = Reactor_cow * 10000 + Actor_cow)

# feed filter
# replacement_THI <- all_replacement_THI %>% 
#   mutate(pair1 = Actor_cow * 10000 + Reactor_cow) %>%
#   mutate(pair2 = Reactor_cow * 10000 + Actor_cow) %>%
#   filter(Bin_type == 'Feed') %>% 
#   filter(pair1 %in% pairs | pair2 %in% pairs)

# feed <- replacement_THI %>% filter(Bin_type == 'csjnskcj')

# for (row in 1:nrow(water_high_groupby)) {
#   i <-  water_high_groupby[row, "replacements"]
#   r <- replacement_THI %>% 
#       filter(pair1 == water_high_groupby[row, "pair1"]$pair1 | pair2 == water_high_groupby[row, "pair2"]$pair2)
#   result <- r[sample(1:i$replacements),]
#   feed <- rbind(feed, result)
#   print(row)
# }
# 
# write.csv(feed, '../data/filtered_feed.csv')



```

```{r elo loop, message=TRUE}
water_feed_cor = 0;
water_feed_high_cor = 0;
water_feed_low_cor = 0;
water_high_low_cor = 0;
feed_high_low_cor = 0;
num <- 1
for (i in 1:num) {
  # feed bins overall
  sub_repl <- feed[sample(1:nrow(water)),]
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores
  elo_res_feed=elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  #water bins overall
  sub_repl <- water
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores 
  elo_res_water = elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  #feed bins high THI
  sub_repl <- feed_high[sample(1:nrow(water_high)),]
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores
  elo_res_feed_high_THI=elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  #feed bins low THI
  sub_repl <- feed_low[sample(1:nrow(water_high)),]
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores
  elo_res_feed_low_THI=elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  #water bins high THI
  sub_repl <- water_high
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores 
  elo_res_water_high_THI=elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  #water bins low THI
  sub_repl <- water_low[sample(1:nrow(water_high)),]
  sub_repl$index <- seq(1, nrow(sub_repl))
  ## Order replacements
  elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
  colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
  elo_repl_list$index <- seq(1, nrow(elo_repl_list))
  #Calculate the ELO scores 
  elo_res_water_low_THI=elo.seq(winner=as.character(elo_repl_list$winner),
                             loser=as.character(elo_repl_list$loser),
                             Date=elo_repl_list$date,
                             k=20,
                             #presence=presence_comb,
                             runcheck = F,
                             progressbar=T)
  
  last_elo_water <- extract_elo(elo_res_water)
  last_elo_feed <- extract_elo(elo_res_feed)
  last_elo_water_low_THI <- extract_elo(elo_res_water_low_THI)
  last_elo_water_high_THI <- extract_elo(elo_res_water_high_THI)
  last_elo_feed_low_THI <- extract_elo(elo_res_feed_low_THI)
  last_elo_feed_high_THI <- extract_elo(elo_res_feed_high_THI)
  
  cows <- Reduce(intersect, list(names(last_elo_water),names(last_elo_feed)))
  sorted_last_elo_feed <- last_elo_feed[sort(cows)]
  sorted_last_elo_water <- last_elo_water[sort(cows)]
  water_feed_cor <- water_feed_cor + cor.test(x=unname(sorted_last_elo_water), y=unname(sorted_last_elo_feed), method = 'spearman')$estimate
  
  low_THI_cows <- Reduce(intersect, list(names(last_elo_water_low_THI),names(last_elo_feed_low_THI)))
  high_THI_cows <- Reduce(intersect, list(names(last_elo_water_high_THI),names(last_elo_feed_high_THI)))
  water_cows <- Reduce(intersect, list(names(last_elo_water_low_THI),names(last_elo_water_high_THI)))
  feed_cows <- Reduce(intersect, list(names(last_elo_feed_low_THI),names(last_elo_feed_high_THI)))
  sorted_last_elo_feed_low_THI <- last_elo_feed_low_THI[sort(low_THI_cows)]
  sorted_last_elo_feed_high_THI <- last_elo_feed_high_THI[sort(high_THI_cows)]
  sorted_last_elo_water_low_THI <- last_elo_water_low_THI[sort(low_THI_cows)]
  sorted_last_elo_water_high_THI <- last_elo_water_high_THI[sort(high_THI_cows)]
  
  water_feed_high_cor <- water_feed_high_cor + cor.test(x=unname(sorted_last_elo_feed_high_THI), y=unname(sorted_last_elo_water_high_THI), method = 'spearman')$estimate
  water_feed_low_cor <- water_feed_low_cor + cor.test(x=unname(sorted_last_elo_feed_low_THI), y=unname(sorted_last_elo_water_low_THI), method = 'spearman')$estimate
  water_high_low_cor <- water_high_low_cor + cor.test(x=unname(last_elo_water_low_THI[sort(water_cows)]), y=unname(last_elo_water_high_THI[sort(water_cows)]), method = 'spearman')$estimate
  feed_high_low_cor <- feed_high_low_cor + cor.test(x=unname(last_elo_feed_low_THI[sort(feed_cows)]), y=unname(last_elo_feed_high_THI[sort(feed_cows)]), method = 'spearman')$estimate
}

water_feed_cor/num
water_feed_high_cor/num
water_feed_low_cor/num
water_high_low_cor/num
feed_high_low_cor/num
```

#### Elo score for Feed bin (Overall)
```{r elo score feed, message=TRUE}
# read.csv('../data/replacement_with_THI.csv')
# replacement_THI <- read.csv('../data/filtered_feed.csv')
# sub_repl <- replacement_THI %>% drop_na()
#shuffle the rows
sub_repl <- feed[sample(1:nrow(water)),]
sub_repl$index <- seq(1, nrow(sub_repl))
## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))
#Calculate the ELO scores
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)
saveRDS(elo_package_result, "../data/elo_res_feed.rds")
logtable_feed <- elo_package_result[["logtable"]]
logtable_feed$index <- seq(1, nrow(logtable_feed))
logtable2_feed <- logtable_feed[, c("winner","loser", "Apost","Bpost", "index")]
logtable3_feed <- merge(logtable2_feed, elo_repl_list)
logtable3_feed$bin <- NULL
write.csv(logtable_feed, '../data/logtable_feed.csv')
write.csv(logtable2_feed, '../data/logtable2_feed.csv')
write.csv(logtable3_feed, '../data/logtable3_feed.csv')
```

#### Elo score for Water bin (Overall)
```{r elo score water, message=TRUE}
# replacement_THI <- read.csv('../data/replacement_with_THI.csv')
sub_repl <- water
#shuffle the rows
#sub_repl <- sub_repl[sample(1:nrow(sub_repl)),]
sub_repl$index <- seq(1, nrow(sub_repl))
## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))
#Calculate the ELO scores 
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)
logtable_water <- elo_package_result[["logtable"]]
logtable_water$index <- seq(1, nrow(logtable_water))
logtable2_water <- logtable_water[, c("winner","loser", "Apost","Bpost", "index")]
logtable3_water <- merge(logtable2_water, elo_repl_list)
logtable3_water$bin <- NULL
saveRDS(elo_package_result, '../data/elo_res_water.rds')
write.csv(logtable_water, '../data/logtable_water.csv')
write.csv(logtable2_water, '../data/logtable2_water.csv')
write.csv(logtable3_water, '../data/logtable3_water.csv')
```


#### Elo score for Feed bin (high THI)(THI_max >= 65)
```{r elo score feed, message=TRUE}
# sub_repl <- replacement_THI %>% filter(Bin_type == 'Feed', THI_max >= 70)
#shuffle the rows
sub_repl <- feed_high[sample(1:nrow(water_high)),]
sub_repl$index <- seq(1, nrow(sub_repl))

## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))

#Calculate the ELO scores
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)

saveRDS(elo_package_result, "../data/elo_res_feed_high_THI.rds")

logtable_feed <- elo_package_result[["logtable"]]
logtable_feed$index <- seq(1, nrow(logtable_feed))

logtable2_feed <- logtable_feed[, c("winner","loser", "Apost","Bpost", "index")]

logtable3_feed <- merge(logtable2_feed, elo_repl_list)
logtable3_feed$bin <- NULL

write.csv(logtable_feed, '../data/logtable_feed_high_THI.csv')
write.csv(logtable2_feed, '../data/logtable2_feed_high_THI.csv')
write.csv(logtable3_feed, '../data/logtable3_feed_high_THI.csv')
```

#### Elo score for Feed bin (low THI) (THI < 65)
```{r elo score feed, message=TRUE}
# sub_repl <- replacement_THI %>% filter(Bin_type == 'Feed', THI_max < 70)
#shuffle the rows
sub_repl <- feed_low[sample(1:nrow(water_high)),]
sub_repl$index <- seq(1, nrow(sub_repl))

## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))

#Calculate the ELO scores
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)

saveRDS(elo_package_result, "../data/elo_res_feed_low_THI.rds")

logtable_feed <- elo_package_result[["logtable"]]
logtable_feed$index <- seq(1, nrow(logtable_feed))

logtable2_feed <- logtable_feed[, c("winner","loser", "Apost","Bpost", "index")]

logtable3_feed <- merge(logtable2_feed, elo_repl_list)
logtable3_feed$bin <- NULL

write.csv(logtable_feed, '../data/logtable_feed_low_THI.csv')
write.csv(logtable2_feed, '../data/logtable2_feed_low_THI.csv')
write.csv(logtable3_feed, '../data/logtable3_feed_low_THI.csv')
```

#### Elo score for Water bin (high THI)(THI_max >= 65)
```{r elo score water, message=TRUE}
# sub_repl <- replacement_THI %>% filter(Bin_type == 'Water', THI_max >= 70)
#shuffle the rows
sub_repl <- water_high
sub_repl$index <- seq(1, nrow(sub_repl))

## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))


#Calculate the ELO scores 
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)

logtable_water <- elo_package_result[["logtable"]]
logtable_water$index <- seq(1, nrow(logtable_water))

logtable2_water <- logtable_water[, c("winner","loser", "Apost","Bpost", "index")]

logtable3_water <- merge(logtable2_water, elo_repl_list)
logtable3_water$bin <- NULL

saveRDS(elo_package_result, '../data/elo_res_water_high_THI.rds')
write.csv(logtable_water, '../data/logtable_water_high_THI.csv')
write.csv(logtable2_water, '../data/logtable2_water_high_THI.csv')
write.csv(logtable3_water, '../data/logtable3_water_high_THI.csv')
```

#### Elo score for Water bin (low THI)(THI_max < 65)
```{r elo score water, message=TRUE}
# sub_repl <- replacement_THI %>% filter(Bin_type == 'Water', THI_max < 70)
#shuffle the rows
sub_repl <- water_low[sample(1:nrow(water_high)),]
sub_repl$index <- seq(1, nrow(sub_repl))

## Order replacements
elo_repl_list <- sub_repl[order(sub_repl$hour),c("Actor_cow", "Reactor_cow","hour", "Bin", "date")]
colnames(elo_repl_list)=c("winner","loser","time","bin", "date")
elo_repl_list$index <- seq(1, nrow(elo_repl_list))


#Calculate the ELO scores 
elo_package_result=elo.seq(winner=as.character(elo_repl_list$winner),
                           loser=as.character(elo_repl_list$loser),
                           Date=elo_repl_list$date,
                           k=20,
                           #presence=presence_comb,
                           runcheck = F,
                           progressbar=T)

logtable_water <- elo_package_result[["logtable"]]
logtable_water$index <- seq(1, nrow(logtable_water))

logtable2_water <- logtable_water[, c("winner","loser", "Apost","Bpost", "index")]

logtable3_water <- merge(logtable2_water, elo_repl_list)
logtable3_water$bin <- NULL

saveRDS(elo_package_result, '../data/elo_res_water_low_THI.rds')
write.csv(logtable_water, '../data/logtable_water_low_THI.csv')
write.csv(logtable2_water, '../data/logtable2_water_low_THI.csv')
write.csv(logtable3_water, '../data/logtable3_water_low_THI.csv')
```
### Load tidy datasets
Now the data is tidied up, This should be run everytime, we start to run this file.
```{r load-logtables, message=TRUE}
# read.csv('../data/replacement_with_THI.csv')
# logtable3_water_high_THI <- read.csv('../data/logtable3_water_high_THI.csv')
# logtable3_water_low_THI <- read.csv('../data/logtable3_water_low_THI.csv')
# logtable3_feed_high_THI <- read.csv('../data/logtable3_feed_high_THI.csv')
# logtable3_feed_low_THI <- read.csv('../data/logtable3_feed_low_THI.csv')
# elo_res_water_low_THI <- readRDS('../data/elo_res_water_low_THI.rds')
# elo_res_water_high_THI <- readRDS('../data/elo_res_water_high_THI.rds')
# elo_res_feed_low_THI <- readRDS('../data/elo_res_feed_low_THI.rds')
# elo_res_feed_high_THI <- readRDS('../data/elo_res_feed_high_THI.rds')
```

Generate/merge log table with the original dataset.
```{r generate-analysis-table, message=TRUE}
# replacement_THI_water <- replacement_THI %>% filter(Bin_type == 'Water')
# replacement_THI_water$index <-seq(1, nrow(replacement_THI_water))
# selected_repcement_THI_water <- replacement_THI_water %>% select(Bin, Bin_type, Bout_interval, THI_mean, THI_max, THI_min, index)
# cowrep_water <- selected_repcement_THI_water %>%
#           inner_join(logtable3_water, by = c("index"))
# 
# replacement_THI_feed <- replacement_THI %>% filter(Bin_type == 'Feed')
# replacement_THI_feed$index <-seq(1, nrow(replacement_THI_feed))
# selected_repcement_THI_feed <- replacement_THI_feed %>% select(Bin, Bin_type, Bout_interval, THI_mean, THI_max, THI_min, index)
# cowrep_feed <- selected_repcement_THI_feed %>%
#           inner_join(logtable3_feed, by = c("index"))
```

## 2. EDA

```{r EDA-1, message=TRUE}
# check assumption (1) before we separate them into high/low THI
# 18 * length(unique(c(unique(replacement_THI_water$Actor_cow), unique(replacement_THI_water$Reactor_cow)))) <= nrow(replacement_THI_water)
# 18 * length(unique(c(unique(replacement_THI_feed$Actor_cow), unique(replacement_THI_feed$Reactor_cow)))) <= nrow(replacement_THI_feed)
# 
# 
# # check assumption (1) after we separate them into high/low THI
# cowrep_THI_high_water <- cowrep_water %>% filter(THI_max >= 65)
# cowrep_THI_low_water<- cowrep_water %>% filter(THI_max < 65)
# cowrep_THI_high_feed <- cowrep_feed %>% filter(THI_max >= 65)
# cowrep_THI_low_feed <- cowrep_feed %>% filter(THI_max < 65)
# 
# 
# 16 * length(unique(c(unique(cowrep_THI_high_water$winner), unique(cowrep_THI_high_water$loser)))) <= nrow(cowrep_THI_high_water)
# 18 * length(unique(c(unique(cowrep_THI_low_water$winner), unique(cowrep_THI_low_water$loser)))) <= nrow(cowrep_THI_low_water)
# 18 * length(unique(c(unique(cowrep_THI_high_feed$winner), unique(cowrep_THI_high_feed$loser)))) <= nrow(cowrep_THI_high_feed)
# 18 * length(unique(c(unique(cowrep_THI_low_feed$winner), unique(cowrep_THI_low_feed$loser)))) <= nrow(cowrep_THI_low_feed)
# 
# 
# cowrep_THI_high_water %>% distinct(winner, loser)
# cowrep_THI_high_feed %>% distinct(winner, loser)
# ```
# ```{r EDA-2, message=FALSE}
# group_THI_List <- list(cowrep_THI_high_water,
#                        cowrep_THI_low_water,
#                        cowrep_THI_high_feed,
#                        cowrep_THI_low_feed)
# 
# lapply(group_THI_List, check_consec_dates)
# 
# check_consec_dates <- 
#   function(group_THI) {
#   group_THI <- group_THI[order(as.Date(group_THI$date, format="%Y-%m-%d")),]
#   group_THI$consecutive <- c(NA,(diff(as.Date(group_THI$date))))
#   group_THI <- group_THI %>% filter(consecutive > 1)
#   return(group_THI)
#   }
```

```{r EDA-3, message=FALSE}
# check_consec_dates(cowrep_feed)
# check_consec_dates(cowrep_water)
```

```{r EDA-4, message=FALSE}
##optimize k
#ores_water <- optimizek(elo_res_water, krange = c(1, 100), resolution = 100)
#ores_water$best

#ores_feed <- optimizek(elo_res_feed, krange = c(1, 100), resolution = 100)
#ores_feed$best

elo_res_water <- readRDS('../data/elo_res_water.rds')
elo_res_feed <- readRDS('../data/elo_res_feed.rds')
elo_res_water_low_THI <- readRDS('../data/elo_res_water_low_THI.rds')
elo_res_water_high_THI <- readRDS('../data/elo_res_water_high_THI.rds')
elo_res_feed_low_THI <- readRDS('../data/elo_res_feed_low_THI.rds')
elo_res_feed_high_THI <- readRDS('../data/elo_res_feed_high_THI.rds')


# ores_water_low_THI <- optimizek(elo_res_water_low_THI, krange = c(1, 100), resolution = 100)
# ores_water_low_THI$best
# ores_water_high_THI <- optimizek(elo_res_water_high_THI, krange = c(1, 100), resolution = 100)
# ores_water_high_THI$best
# ores_feed_low_THI <- optimizek(elo_res_feed_low_THI, krange = c(1, 100), resolution = 100)
# ores_feed_low_THI$best
# ores_feed_high_THI <- optimizek(elo_res_feed_high_THI, krange = c(1, 100), resolution = 100)
# ores_feed_high_THI$best


last_elo_water <- extract_elo(elo_res_water)
last_elo_feed <- extract_elo(elo_res_feed)
last_elo_water_low_THI <- extract_elo(elo_res_water_low_THI)
last_elo_water_high_THI <- extract_elo(elo_res_water_high_THI)
last_elo_feed_low_THI <- extract_elo(elo_res_feed_low_THI)
last_elo_feed_high_THI <- extract_elo(elo_res_feed_high_THI)

cows <- Reduce(intersect, list(names(last_elo_water),names(last_elo_feed)))
sorted_last_elo_feed <- last_elo_feed[sort(cows)]
sorted_last_elo_water <- last_elo_water[sort(cows)]
cor.test(x=unname(sorted_last_elo_water), y=unname(sorted_last_elo_feed), method = 'spearman')


low_THI_cows <- Reduce(intersect, list(names(last_elo_water_low_THI),names(last_elo_feed_low_THI)))
high_THI_cows <- Reduce(intersect, list(names(last_elo_water_high_THI),names(last_elo_feed_high_THI)))
water_cows <- Reduce(intersect, list(names(last_elo_water_low_THI),names(last_elo_water_high_THI)))
feed_cows <- Reduce(intersect, list(names(last_elo_feed_low_THI),names(last_elo_feed_high_THI)))
sorted_last_elo_feed_low_THI <- last_elo_feed_low_THI[sort(low_THI_cows)]
sorted_last_elo_feed_high_THI <- last_elo_feed_high_THI[sort(high_THI_cows)]
sorted_last_elo_water_low_THI <- last_elo_water_low_THI[sort(low_THI_cows)]
sorted_last_elo_water_high_THI <- last_elo_water_high_THI[sort(high_THI_cows)]




cor.test(x=unname(sorted_last_elo_feed_high_THI), y=unname(sorted_last_elo_water_high_THI), method = 'spearman')
cor.test(x=unname(sorted_last_elo_feed_low_THI), y=unname(sorted_last_elo_water_low_THI), method = 'spearman')
cor.test(x=unname(last_elo_water_low_THI[sort(water_cows)]), y=unname(last_elo_water_high_THI[sort(water_cows)]), method = 'spearman')
cor.test(x=unname(last_elo_feed_low_THI[sort(feed_cows)]), y=unname(last_elo_feed_high_THI[sort(feed_cows)]), method = 'spearman')
```
```{r elo-steepness, message=FALSE}
# library(EloSteepness)
# water_high_matrix <- creatematrix(elo_res_water_high_THI)
# water_low_matrix <- creatematrix(elo_res_water_low_THI)
# feed_high_matrix <- creatematrix(elo_res_feed_high_THI)
# feed_low_matrix <- creatematrix(elo_res_feed_low_THI)
# 
# # using small numbers for iterations etc to speed up running time
# water_high_res <- elo_steepness_from_matrix(water_high_matrix, cores = 2,iter = 50, warmup = 25,refresh = 0, chains = 4, seed = 1)
# plot_scores(water_high_res)
# 
# water_low_res <- elo_steepness_from_matrix(water_low_matrix, cores = 2,iter = 50, warmup = 25,refresh = 0, chains = 4, seed = 1)
# plot_scores(water_low_res)
# 
# feed_high_res <- elo_steepness_from_matrix(feed_high_matrix, cores = 2,iter = 50, warmup = 25,refresh = 0, chains = 4, seed = 1)
# plot_scores(feed_high_res)
# 
# feed_low_res <- elo_steepness_from_matrix(feed_low_matrix, cores = 2,iter = 50, warmup = 25,refresh = 0, chains = 4, seed = 1)
# plot_scores(feed_low_res)


```
